Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2022-10-03 src/tex/ms.tex
text changes for referee second round

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\cosmic}{\texttt{COSMIC}}
\newcommand{\legwork}{\texttt{LEGWORK}}
\usepackage{xcolor}
\usepackage{showyourwork}
\newcommand{\snotes}[1]{\textcolor{purple}{#1}}
\newcommand{\kb}[1]{\textcolor{teal}{#1}}

@@

% Abstract with filler text
\begin{abstract}
Short-period double white dwarf (DWD) binaries will be the most prolific source of gravitational waves (GWs) for the Laser Interferometer Space Antenna (LISA). DWDs with GW frequencies below $\sim1\,\rm{mHz}$ will be the dominant contributor to a stochastic foreground caused by overlapping GW signals. Population modeling of Galactic DWDs typically assumes a binary fraction of $50\%$ \textbf{and a log-uniform Zero Age Main Sequence (ZAMS) orbital period distribution}. However, recent observations have shown that the binary fraction of close, solar-type stars exhibits a strong anti-correlation with metallicity \textbf{which modulates the ZAMS orbital period distribution below $10^4$ days}. In this study we perform the first simulation of the Galactic DWD population observable by LISA which incorporates an empirically-derived metallicity-dependent binary fraction, \textbf{using the binary population synthesis suite \cosmic\ and a metallicity-dependent star formation history}. We compare two models: one which assumes a metallicity-dependent binary fraction, and one with a binary fraction of 50\%. \textbf{We repeat our analysis for three different assumptions for Roche-lobe overflow interactions.} We find that while metallicity impacts the evolution and intrinsic properties of our simulated DWD progenitor binaries, the LISA-resolvable populations of the two models remain roughly indistinguishable. However, the size of the total Galactic DWD population orbiting in the LISA frequency band is reduced by more than half when accounting for a metallicity-dependent binary fraction \textbf{for two of our four variations, which also lowers the effective foreground. The LISA population remains unchanged in number for two variations, highlighting the sensitivity of the population to binary evolution prescriptions.}
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the stars which are born with a binary companion, many will undergo interactions which bring the two stars closer together, eventually forming a close double white dwarf (DWD). Close DWDs, with orbital periods shorter than $\sim2.75\,\rm{hr}$ are the largest source by number of mHz gravitational waves (GWs) in the Galaxy \citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space Antenna (LISA) is expected to resolve at least $10^4$ individual DWD binaries in the Milky Way and will also observe GW emission from the entire Galactic DWD population through the unresolved foreground created by overlapping signals at sub-mHz frequencies \citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, Korol2017, Lamberts2019, Breivik2020a}. The resolved population will enable the study of several important aspects of binary evolution like the strength of tides \citep{Valsecchi2012}, the stability of mass transfer in DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, Sepinsky2014, Kremer2015}\textbf{, and the separation distribution of close DWDs \citep{Korol2021}} as well as provide a probe of Galactic structure \citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and strength of the Galactic DWD foreground can also be used as a tool to study the structure of the Milky Way \citep{Benacquista2006, Breivik2020b}.

\textbf{The transition between individually resolved DWDs and the confusion limited, or unresolved, DWD foreground is expected to occur near $\sim1\,\rm{mHz}$ frequencies \citep[e.g. ][]{Ruiter2010}. In the confusion limited regime, more than one binary radiates GWs in each LISA frequency bin, thus creating a superposition of signals which are unable to be disentangled. There are multiple parameters which shape the Galactic DWD foreground. Assuming the evolution of each DWD system is driven solely by GW emission, the frequency derivative scales proportionally to $f_{\rm{GW}}^{11/3}$, causing a pileup of DWDs at lower frequencies \citep[e.g.][]{Breivik2020b}. The spatial density of DWDs in the Galaxy which defines the distance to each binary also impacts the foreground amplitude, since GW strain scales inversely with distance. Star formation history assumptions combine these two effects by assigning ages and positions to each DWD in the population which determine the present day orbital (and thus GW) frequencies as well as distances to each source.}

When viewed strictly as a source of noise, the unresolved Galactic DWD foreground is the dominant noise source for LISA in the sub-mHz part of LISA's frequency range. This extra noise above the detector noise floor affects the detection of all other LISA sources including extreme mass ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, Moore2017}, merging black holes with masses between $10^4$--$10^7$ M$_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, Caldwell2019}. For sources which have signals buried by the Galactic DWD population, the foreground must be carefully analyzed and subtracted \citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number of resolved DWDs and the height of the unresolved DWD foreground are a direct consequence of the number of DWD progenitors which form and evolve over the Milky Way's history.

While the binary fraction remains approximately constant across a large metallicity range ($-1.5 \leq$ [Fe/H] $< 0.5$) for wide binaries, close OB stars, and the stellar Initial Mass Function (IMF) \citep{Moe2017, Moe2019}, the binary fraction for solar-type star systems with orbital period $P_{\text{orb}}\leq 10^4$ days (separation $a \leq 10$ AU) shows a strong anti-correlation with metallicity \citep[e.g.][]{Badenes2018, Moe2019, Mazzola2020, Price-Whelan2020}. Because close DWDs are the remnants of close, solar-type binary stars, this anti-correlation plays an important role in the formation, evolution, and characteristics of the DWD population that LISA will observe.

@@

In Section \ref{sec:simulations} we discuss our assumptions used in simulating DWD populations and detail the process to produce present-day synthetic Milky-Way-like galaxies. In Section \ref{sec:LISA_obs} we review the derivation of LISA detectability for circular DWD populations at mHz frequencies. In Section \ref{sec:results} we detail results showing how a metallicity-dependent binary fraction affects the formation and evolution of DWD populations assuming \textbf{four sets of} of binary evolution assumptions. In Section \ref{sec:LISA_met} we detail the metallicity dependence of the LISA DWD foreground and resolved population. Finally, \textbf{we consider how a metallicity-dependent close binary fraction impacts the height} of the Galactic DWD foreground in LISA for \textbf{different binary evolution assumptions} in Section \ref{sec:model_compare} and conclude in Section \ref{sec:conclusions}.

\section{Simulating a Galactic DWD population}\label{sec:simulations}
In this section we describe the setup of our DWD simulations using the binary population synthesis suite \cosmic, and the process to scale these simulations to create Milky Way-like galaxies using the star formation history of galaxy \textbf{\textbf{m12i}} in the Latte suite of the FIRE-2 simulations \citep{Wetzel2016, Hopkins2018} and stellar positions assigned according to the the Ananke framework \citep{Sanderson2020}.

\subsection{Metallicity-dependent binary fraction}\label{subsec:metbinfrac}
\textbf{The close binary fraction in the Galaxy has been empirically shown to depend on metallicity. This is manifested through an explicit dependence of the companion frequency on orbital period \citep{Moe2021}. When anchored to a log-normal orbital period distribution, the \emph{frequency} of solar-type binary companions is skewed to shorter orbital periods. The companion frequency differs from the binary fraction: the companion frequency is the probability of a star having a companion at a given orbital period, and the integration of this probability over all periods results in the binary fraction of the population. The metallicity dependence of the companion frequency is observed only below P$_{\rm{orb}}=10^6$ days, above which there are not sufficient observational data of binary systems to infer a metallicity dependence.}

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/met_dep_porb.pdf}
    \caption{The distribution of the companion frequency f$_{\rm{logP}}$ for solar-type stars is dependent on both ZAMS orbital period and metallicity. Each curve is coloured by metallicity, with the darkest curve corresponding to the lowest-metallicity bin in our simulations at [Fe/H]$=-2.3$ and the lightest curve corresponding to our highest-metallicity bin at [Fe/H]$=0.18$. Above $\log_{10}(P_{\rm{orb}}/\rm{day})\sim 4$, the companion frequency distribution is the same for all metallicity bins as no identified metallicity dependence at those periods exists in current literature.}
    \label{fig:met_dep_porb}
    \script{met_dep_porb.py}
\end{figure}

\textbf{Figure \ref{fig:met_dep_porb} shows the companion frequency f$_{\rm{logP}}$ for solar-type stars as a function of the logarithm of orbital period, mirroring the inner binaries curve of Figure 2 in \citet{Moe2021} and scaling to an overall multiplicity frequency of 0.67 (equivalent to an integrated binary fraction of 50\%) at solar metallicity. To obtain the normalization of the companion frequency for close binaries (P$_{\rm{orb}}<10^4$),} we fit the results presented in \citet{Moe2019} using linear regression to obtain a piecewise relation between the metallicity, [Fe/H] and \text{close} binary fraction $f_{\rm{b}}$ as

@@

\noindent where we convert between [Fe/H] and metallicity Z, assuming all stars have solar abundance such that
\begin{equation}
\label{eq:metallicity}
    [\rm{Fe/H}] = \log_{10}\Big(\frac{Z}{Z_{\odot}}\Big),
\end{equation}
\noindent where we assume $Z_{\odot}=0.02$.

\textbf{We count all higher-order companions as secondary, or binary, components. We show curves for all 15 of our metallicity bins, increasing in metallicity from darkest to lightest curve colour. The lowest metallicity ([Fe/H]=-2.3) shows the highest propensity for short-period binaries, in contrast to higher metallicity systems which have a frequency distribution with a relatively constant companion frequency for orbital periods between $10^4$ and $10^6$ days.}

\textbf{While we simulate binaries across the entire orbital period distribution, the close binary fraction, which exhibits the discussed metallicity dependence, will be that which impacts the LISA population since systems with $\log_{10}(\rm{P}_{\rm{orb}}/$day)$>4$ will be unlikely to evolve enough to enter the LISA band by present day.}

@@

The masses and orbital periods at the formation of the second-formed WD span a wide range depending on the WD binary component types, thus we consider four DWD combinations: two helium WDs (He + He), a carbon-oxygen WD orbiting a helium WD (CO + He), two carbon-oxygen WDs (CO + CO), and an oxygen-neon orbiting a helium, carbon-oxygen, or oxygen-neon WD (ONe + X). For each DWD type we simulate a grid of $15$ metallicities spaced uniformly in $\log_{10}(Z)$ between $Z=10^{-4}$ to $0.03$, to account for the limits of the \citet{Hurley2000} stellar evolution tracks employed in \cosmic. This results in a total of $60$ populations across all DWD types and metallicities \textbf{for each set of model assumptions}. The output of \cosmic\ contains information limited to intrinsic binary properties like mass and orbital period. External parameters like Galactic position and orientation are assigned in a post-processing scheme which uses metallicity-dependent positions and ages from the Ananke framework of galaxy \textbf{m12i} from the Latte Suite of the FIRE-2 simulations (see Section~\ref{sec:FIRE} for details).

We assume that the Zero Age Main Sequence (ZAMS) masses, orbital periods and eccentricities for each binary are independently distributed. We choose primary masses following \citet{Kroupa2001}, a flat mass ratio distribution \citep{Mazeh1992, Goldberg1994} and a uniform eccentricity distribution following \cite{Geller2019}. \textbf{For our simulations which follow a metallicity-dependent binary fraction (model FZ), a skewed log-normal orbital period distribution is sampled in accordance with the metallicity-dependent companion frequency of \citet{Moe2021} as discussed in Section \ref{subsec:metbinfrac}. For our simulations which assume a constant 50\% binary fraction (model F50), we use a log-uniform period distribution following Opik's Law similar to previous studies \citep[e.g.][]{Nelemans2001a, Toonen2012, Korol2017, Lamberts2019}. We assume a $100\%$ binary fraction in the F50 COSMIC simulations to reduce computation time and scale the simulations to a constant $50\%$ binary fraction in a post-processing scheme.} We initialize all binaries with the same evolution time of $13.7\,\rm{Gyr}$ to capture all potential evolution within a Hubble time. \textbf{Although we simulate binaries across the full orbital period distribution, we discard DWD binaries which form with separations $a\geq 1000$ R$_\odot$ during post-processing, since these are unlikely to evolve into the LISA band by present day.}

\textbf{We consider a fiducial set of assumptions which} follow the \texttt{COSMIC} defaults described in \citet{Breivik2020a} except for the treatment of Roche-lobe overflow (RLO). The stability of RLO mass transfer is determined using critical mass ratios resulting from radius-mass exponents \citep{Webbink1985, Hurley2002}, where the critical mass ratio is defined as the ratio of the donor to accretor mass. We assume critical mass ratios following \citet{Claeys2014} which reduce the standard critical mass ratio assumptions from \citet{Hurley2002} for main sequence (MS) donors by $\sim50\%$ from $3$ to $1.6$ based on the models of \citet{deMink2007} and treat WD accretors separately following the models of \citet{Soberman1997}. We increase the mass loss rate from the donor following Equation 11 of \citet{Claeys2014}. The amount of mass lost during RLO from the donor is limited by the overflow factor of the donor radius to its Roche radius following \citet{Hurley2002}. The amount of mass accepted by the accretor is limited to $10$ times the accretor's mass divided by the accretor's thermal timescale. Finally, for RLO mass loss which becomes unstable and leads to common envelope evolution (CEE) we assume that the donor's binding energy is calculated according to the fits detailed in Appendix B of \citet{Claeys2014} and that orbital energy is deposited with $100\%$ efficiency into unbinding the common envelope ($\alpha=1$).

\textbf{The LISA-detectable DWD populations that result from our simulations may vary depending on the assumptions made regarding binary interactions. In order to explore the range of results, we complete this study for three binary evolution parameter variations on top of our fiducial set of assumptions. For each variation, we consider models FZ and F50 as done in the fiducial case described above. In variation $q3$, we vary the assumption for the critical mass ratios at which a RLO interaction remains stable or becomes unstable from our fiducial assumptions such that the critical mass ratio is increased to $3.0$ and thus allows stable mass transfer for more massive RLO donors. In variations $\alpha25$ and $\alpha5$, we modify the common envelope ejection efficiency to be either much less ($\alpha=0.25$) or more ($\alpha=5$) than in our fiducial assumption ($\alpha=1$) to capture the range of possible CE ejection efficiencies quoted in the literature  \citep[e.g.][]{Zorotovic2010, Fragos2019}. Larger common envelope ejection efficiencies lead to wider post-CE separations, while smaller ejection efficiencies either lead to closer post-CE separations or stellar mergers where the envelope ejection fails. Between all variations, all other prescriptions for binary evolution remain identical besides the one varied parameter. We delve in depth into the fiducial results in each section, and then give a brief overview of how the results change for each parameter variation.}

@@

The \textbf{m12i} galaxy provides particle mass resolution of 7070 M$_\odot$ per star particle. Each star particle has an associated metallicity, position, and age, which is combined with the output of \cosmic\ to assign DWDs to each star particle by matching its metallicity to our \cosmic\ metallicity grid. The positions of each DWD are assigned using the Ananke framework since multiple DWD binaries can form within a single star particle. Specifically, we use an epanechnikov kernel where the kernel size is inversely proportional to the local density to assign the radial component of spherically symmetric offsets from the center of each star particle following \citet{Sanderson2020}.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close solar-type binaries with P $< 10^4$ days (black) plotted against the logarithm of Z/Z$_\odot$, with scatter points denoting the location of the metallicity grid used in our \cosmic\ simulations. The secondary axis shows the amount of mass in stars formed within each metallicity bin from galaxy \textbf{m12i} of the ``Latte" suite in the FIRE-2 simulations as a red histogram. The amount of stellar mass formed with super-solar metallicities dominates the distribution. Note that the primary axis shows $f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of mass formed in log-scale. The opposing trends of these two distributions compete throughout this study.}
    \label{fig:sfh_vs_fb}
    \script{sfh_vs_fb.py}
\end{figure}

The metallicity-dependent \textbf{close} binary fraction \textbf{for solar-type binaries (P $< 10^4$ days), or simply the close binary fraction from here on, } is shown in black in Figure \ref{fig:sfh_vs_fb} along with the mass in star particles from galaxy \textbf{m12i}, shown in red, as a function of our metallicity grid. The \textbf{close} binary fraction, $f_{\rm{b}}$, drops drastically across metallicity while the mass formed in \textbf{m12i} increases significantly. These two opposing trends compete throughout this study along with the impact of metallicity on single star evolution to form the final numerical distribution of systems in our DWD populations.

>>>>

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\cosmic}{\texttt{COSMIC}}
\newcommand{\legwork}{\texttt{LEGWORK}}
\usepackage{xcolor}
\usepackage{showyourwork}
\newcommand{\snotes}[1]{\textcolor{purple}{#1}}
\newcommand{\kb}[1]{\textcolor{teal}{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

@@

% Abstract with filler text
\begin{abstract}
Short-period double white dwarf (DWD) binaries will be the most prolific 
source of gravitational waves (GWs) for the Laser Interferometer Space 
Antenna (LISA). DWDs with GW frequencies below $\sim1\,\rm{mHz}$ will be 
the dominant contributor to a stochastic foreground caused by overlapping 
GW signals. Population modeling of Galactic DWDs typically assumes a 
binary fraction of $50\%$ and a log-uniform Zero Age Main Sequence (ZAMS) 
orbital period distribution. However, recent observations have shown that 
the binary fraction of close, solar-type stars exhibits a strong 
anti-correlation with metallicity which modulates the ZAMS orbital period 
distribution below $10^4$ days. In this study we perform the first 
simulation of the Galactic DWD population observable by LISA which 
incorporates an empirically-derived metallicity-dependent binary fraction, 
using the binary population synthesis suite \cosmic\ and a 
metallicity-dependent star formation history. We compare two models: one 
which assumes a metallicity-dependent binary fraction, and one with a 
binary fraction of 50\%. We repeat our analysis for three different 
assumptions for Roche-lobe overflow interactions. We find that while 
metallicity impacts the evolution and intrinsic properties of our 
simulated DWD progenitor binaries, the LISA-resolvable populations of the 
two models remain roughly indistinguishable. However, the size of the 
total Galactic DWD population orbiting in the LISA frequency band is 
reduced by more than half when accounting for a metallicity-dependent 
binary fraction for two of our four variations, which also lowers the 
effective foreground. The LISA population remains unchanged in number for 
two variations, highlighting the sensitivity of the population to binary 
evolution prescriptions.
\end{abstract}

@@

\section{Introduction} \label{sec:intro}
Most stars in the Galaxy will end their lives as white dwarfs. Of the 
stars which are born with a binary companion, many will undergo 
interactions which bring the two stars closer together, eventually forming 
a close double white dwarf (DWD). Close DWDs, with orbital periods shorter 
than $\sim2.75\,\rm{hr}$ are the largest source by number of mHz 
gravitational waves (GWs) in the Galaxy 
\citep[e.g.,][]{LISAMissionProposal}. The Laser Interferometer Space 
Antenna (LISA) is expected to resolve at least $10^4$ individual DWD 
binaries in the Milky Way and will also observe GW emission from the 
entire Galactic DWD population through the unresolved foreground created 
by overlapping signals at sub-mHz frequencies 
\citep[e.g.,][]{Nelemans2001a, Ruiter2010, Nissanke2012, Yu2013, 
Korol2017, Lamberts2019, Breivik2020a}. The resolved population will 
enable the study of several important aspects of binary evolution like the 
strength of tides \citep{Valsecchi2012}, the stability of mass transfer in 
DWD systems \citep[e.g.,][]{Marsh2004, Shen2015, Gokhale2007, 
Sepinsky2014, Kremer2015}, and the separation distribution of close DWDs 
\citep{Korol2021} as well as provide a probe of Galactic structure 
\citep{Korol2019} and the Local Group \citep{Korol2018}. The shape and 
strength of the Galactic DWD foreground can also be used as a tool to 
study the structure of the Milky Way \citep{Benacquista2006, 
Breivik2020b}.

The transition between individually resolved DWDs and the confusion 
limited, or unresolved, DWD foreground is expected to occur near 
$\sim1\,\rm{mHz}$ frequencies \citep[e.g. ][]{Ruiter2010}. In the 
confusion limited regime, more than one binary radiates GWs in each LISA 
frequency bin, thus creating a superposition of signals which are unable 
to be disentangled. There are multiple parameters which shape the Galactic 
DWD foreground. Assuming the evolution of each DWD system is driven solely 
by GW emission, the frequency derivative scales proportionally to 
$f_{\rm{GW}}^{11/3}$, causing a pileup of DWDs at lower frequencies 
\citep[e.g.][]{Breivik2020b}. The spatial density of DWDs in the Galaxy 
which defines the distance to each binary also impacts the foreground 
amplitude, since GW strain scales inversely with distance. Star formation 
history assumptions combine these two effects by assigning ages and 
positions to each DWD in the population which determine the present day 
orbital (and thus GW) frequencies as well as distances to each source.

When viewed strictly as a source of noise, the unresolved Galactic DWD 
foreground is the dominant noise source for LISA in the sub-mHz part of 
LISA's frequency range. This extra noise above the detector noise floor 
affects the detection of all other LISA sources including extreme mass 
ratio inspirals \citep[e.g.,][]{Berti2006, Barack2007, Babak2017, 
Moore2017}, merging black holes with masses between 
$10^4$--$10^7\,M_{\odot}$ \citep[e.g.,][]{Klein2016, Bellovary2019}, and 
cosmological GW backgrounds \citep[e.g.,][]{Bartolo2016, Caprini2016, 
Caldwell2019}. For sources which have signals buried by the Galactic DWD 
population, the foreground must be carefully analyzed and subtracted 
\citep[][]{Adams2014,Cornish2020,Littenberg2020,Boileau2021}. The number 
of resolved DWDs and the height of the unresolved DWD foreground are a 
direct consequence of the number of DWD progenitors which form and evolve 
over the Milky Way's history.

While the binary fraction remains approximately constant across a large 
metallicity range ($-1.5 \leq$ [Fe/H] $< 0.5$) for wide binaries, close OB 
stars, and the stellar Initial Mass Function (IMF) \citep{Moe2017, 
Moe2019}, the binary fraction for solar-type star systems with orbital 
period $P_{\rm{orb}}\leq 10^4$ days (separation $a \leq 10$ AU) shows a 
strong anti-correlation with metallicity \citep[e.g.][]{Badenes2018, 
Moe2019, Mazzola2020, Price-Whelan2020}. Because close DWDs are the 
remnants of close, solar-type binary stars, this anti-correlation plays an 
important role in the formation, evolution, and characteristics of the DWD 
population that LISA will observe.

@@

In Section \ref{sec:simulations} we discuss our assumptions used in 
simulating DWD populations and detail the process to produce present-day 
synthetic Milky-Way-like galaxies. In Section \ref{sec:LISA_obs} we review 
the derivation of LISA detectability for circular DWD populations at mHz 
frequencies. In Section \ref{sec:results} we detail results showing how a 
metallicity-dependent binary fraction affects the formation and evolution 
of DWD populations assuming four sets of of binary evolution assumptions. 
In Section \ref{sec:LISA_met} we detail the metallicity dependence of the 
LISA DWD foreground and resolved population. Finally, we consider how a 
metallicity-dependent close binary fraction impacts the height of the 
Galactic DWD foreground in LISA for different binary evolution assumptions 
in Section \ref{sec:model_compare} and conclude in Section 
\ref{sec:conclusions}.

\section{Simulating a Galactic DWD population}\label{sec:simulations}
In this section we describe the setup of our DWD simulations using the 
binary population synthesis suite \cosmic, and the process to scale these 
simulations to create Milky Way-like galaxies using the star formation 
history of galaxy \textbf{m12i} in the Latte suite of the FIRE-2 
simulations \citep{Wetzel2016, Hopkins2018} and stellar positions assigned 
according to the the Ananke framework \citep{Sanderson2020}.

\subsection{Metallicity-dependent binary 
fraction}\label{subsec:metbinfrac}
The close binary fraction in the Galaxy has been empirically shown to 
depend on metallicity. This is manifested through an explicit dependence 
of the companion frequency on orbital period \citep{Moe2021}. When 
anchored to a log-normal orbital period distribution, the \emph{frequency} 
of solar-type binary companions is skewed to shorter orbital periods. The 
companion frequency differs from the binary fraction: the companion 
frequency is the probability of a star having a companion at a given 
orbital period, and the integration of this probability over all periods 
results in the binary fraction of the population. The metallicity 
dependence of the companion frequency is observed only below 
$P_{\rm{orb}}=10^6$ days, above which there are not sufficient 
observational data of binary systems to infer a metallicity dependence.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/met_dep_porb.pdf}
    \caption{The distribution of the companion frequency f$_{\rm{logP}}$ 
for solar-type stars is dependent on both ZAMS orbital period and 
metallicity. Each curve is coloured by metallicity, with the darkest curve 
corresponding to the lowest-metallicity bin in our simulations at [Fe/H] 
$=-2.3$ and the lightest curve corresponding to our highest-metallicity 
bin at [Fe/H] $=0.18$. Above $\log_{10}(P_{\rm{orb}}/\rm{day})\sim 4$, the 
companion frequency distribution is the same for all metallicity bins as 
no identified metallicity dependence at those periods exists in current 
literature.}
    \label{fig:met_dep_porb}
    \script{met_dep_porb.py}
\end{figure}

Figure \ref{fig:met_dep_porb} shows the companion frequency 
f$_{\rm{logP}}$ for solar-type stars as a function of the logarithm of 
orbital period, mirroring the inner binaries curve of Figure 2 in 
\citet{Moe2021} and scaling to an overall multiplicity frequency of 0.67 
(equivalent to an integrated binary fraction of 50\%) at solar 
metallicity. To obtain the normalization of the companion frequency for 
close binaries ($P_{\rm{orb}}<10^4$), we fit the results presented in 
\citet{Moe2019} using linear regression to obtain a piecewise relation 
between the metallicity, [Fe/H] and \textbf{close} binary fraction 
$f_{\rm{b}}$ as

@@

\noindent where we convert between [Fe/H] and metallicity Z, assuming all 
stars have solar abundance such that
\begin{equation}
\label{eq:metallicity}
    [\rm{Fe/H}] = 
\log_{10}\Big(\frac{\textit{Z}}{\textit{Z}_{\odot}}\Big),
\end{equation}
\noindent where we assume $Z_{\odot}=0.02$.

We count all higher-order companions as secondary, or binary, components. 
We show curves for all 15 of our metallicity bins, increasing in 
metallicity from darkest to lightest curve colour. The lowest metallicity 
($\rm{[Fe/H]}=-2.3$) shows the highest propensity for short-period 
binaries, in contrast to higher metallicity systems which have a frequency 
distribution with a relatively constant companion frequency for orbital 
periods between $10^4$ and $10^6$ days.

While we simulate binaries across the entire orbital period distribution, 
the close binary fraction, which exhibits the discussed metallicity 
dependence, will be that which impacts the LISA population since systems 
with $\log_{10}(P_{\rm{orb}}/$day)$>4$ will be unlikely to evolve enough 
to enter the LISA band by present day.

@@

The masses and orbital periods at the formation of the second-formed WD 
span a wide range depending on the WD binary component types, thus we 
consider four DWD combinations: two helium WDs (He + He), a carbon-oxygen 
WD orbiting a helium WD (CO + He), two carbon-oxygen WDs (CO + CO), and an 
oxygen-neon orbiting a helium, carbon-oxygen, or oxygen-neon WD (ONe + X). 
For each DWD type we simulate a grid of $15$ metallicities spaced 
uniformly in $\log_{10}(Z)$ between $Z=10^{-4}$ to $0.03$, to account for 
the limits of the \citet{Hurley2000} stellar evolution tracks employed in 
\cosmic. This results in a total of $60$ populations across all DWD types 
and metallicities for each set of model assumptions. The output of 
\cosmic\ contains information limited to intrinsic binary properties like 
mass and orbital period. External parameters like Galactic position and 
orientation are assigned in a post-processing scheme which uses 
metallicity-dependent positions and ages from the Ananke framework of 
galaxy \textbf{m12i} from the Latte Suite of the FIRE-2 simulations (see 
Section~\ref{sec:FIRE} for details).

We assume that the Zero Age Main Sequence (ZAMS) masses, orbital periods 
and eccentricities for each binary are independently distributed. We 
choose primary masses following \citet{Kroupa2001}, a flat mass ratio 
distribution \citep{Mazeh1992, Goldberg1994} and a uniform eccentricity 
distribution following \cite{Geller2019}. For our simulations which follow 
a metallicity-dependent binary fraction (model FZ), a skewed log-normal 
orbital period distribution is sampled in accordance with the 
metallicity-dependent companion frequency of \citet{Moe2021} as discussed 
in Section \ref{subsec:metbinfrac}. For our simulations which assume a 
constant 50\% binary fraction (model F50), we use a log-uniform period 
distribution following Opik's Law similar to previous studies 
\citep[e.g.][]{Nelemans2001a, Toonen2012, Korol2017, Lamberts2019}. We 
assume a $100\%$ binary fraction in the F50 COSMIC simulations to reduce 
computation time and scale the simulations to a constant $50\%$ binary 
fraction in a post-processing scheme. We initialize all binaries with the 
same evolution time of $13.7\,\rm{Gyr}$ to capture all potential evolution 
within a Hubble time. Although we simulate binaries across the full 
orbital period distribution, we discard DWD binaries which form with 
separations $a\geq 1000\,R_\odot$ during post-processing, since these are 
unlikely to evolve into the LISA band by present day.

We consider a fiducial set of assumptions which follow the \cosmic\ 
defaults described in \citet{Breivik2020a} except for the treatment of 
Roche-lobe overflow (RLO). The stability of RLO mass transfer is 
determined using critical mass ratios resulting from radius-mass exponents 
\citep{Webbink1985, Hurley2002}, where the critical mass ratio is defined 
as the ratio of the donor to accretor mass. We assume critical mass ratios 
following \citet{Claeys2014} which reduce the standard critical mass ratio 
assumptions from \citet{Hurley2002} for main sequence (MS) donors by 
$\sim50\%$ from $3$ to $1.6$ based on the models of \citet{deMink2007} and 
treat WD accretors separately following the models of 
\citet{Soberman1997}. We increase the mass loss rate from the donor 
following Equation 11 of \citet{Claeys2014}. The amount of mass lost 
during RLO from the donor is limited by the overflow factor of the donor 
radius to its Roche radius following \citet{Hurley2002}. The amount of 
mass accepted by the accretor is limited to $10$ times the accretor's mass 
divided by the accretor's thermal timescale. Finally, for RLO mass loss 
which becomes unstable and leads to common envelope evolution (CEE) we 
assume that the donor's binding energy is calculated according to the fits 
detailed in Appendix B of \citet{Claeys2014} and that orbital energy is 
deposited with $100\%$ efficiency into unbinding the common envelope 
($\alpha=1$).

The LISA-detectable DWD populations that result from our simulations may 
vary depending on the assumptions made regarding binary interactions. In 
order to explore the range of results, we complete this study for three 
binary evolution parameter variations on top of our fiducial set of 
assumptions. For each variation, we consider models FZ and F50 as done in 
the fiducial case described above. In variation $q3$, we vary the 
assumption for the critical mass ratios at which a RLO interaction remains 
stable or becomes unstable from our fiducial assumptions. The critical 
mass ratio $q_c$ is increased to $3.0$ and thus allows stable mass 
transfer for more massive RLO donors. In variations $\alpha25$ and 
$\alpha5$, we modify the common envelope ejection efficiency to be either 
much less ($\alpha=0.25$) or more ($\alpha=5$) than in our fiducial 
assumption ($\alpha=1$) to capture the range of possible CE ejection 
efficiencies quoted in the literature  \citep[e.g.][]{Zorotovic2010, 
Fragos2019}. Larger common envelope ejection efficiencies lead to wider 
post-CE separations, while smaller ejection efficiencies either lead to 
closer post-CE separations or stellar mergers where the envelope ejection 
fails. Between all variations, all other prescriptions for binary 
evolution remain identical besides the one varied parameter. \textbf{A 
summary of all models and variations explored in this work is given in 
Table \ref{tbl:models}.} We delve in depth into the fiducial results in 
each section, and then give a brief overview of how the results change for 
each parameter variation.

\begin{table}[]
    \centering
    \begin{tabular}{|c|c|c|}
       \hline
       model & binary fraction & orbital period distribution \\
       \hline
       FZ & \citet{Moe2019} & \citet{Moe2021}\\
       F50 & 50\% & flat in $\log(P_{\rm{orb}}/\rm{day})$\\
       \hline
       \hline       
       variation & parameter change & binary evolution change \\
       \hline
       fiducial  & none & none \\
       $\alpha25$ & $\alpha=0.25$ & reduced CE efficiency \\
       $\alpha5$  & $\alpha=5$ & increased CE efficiency\\
       $q3$ & $q_c=3$ & increased critical mass ratio \\
       \hline
    \end{tabular}
    \caption{\textbf{Summary of the models and binary evolution assumption 
variation simulations that are explored in this study.}}
    \label{tbl:models}
\end{table}

@@

The \textbf{m12i} galaxy provides particle mass resolution of 7070 
$M_\odot$ per star particle. Each star particle has an associated 
metallicity, position, and age, which is combined with the output of 
\cosmic\ to assign DWDs to each star particle by matching its metallicity 
to our \cosmic\ metallicity grid. The positions of each DWD are assigned 
using the Ananke framework since multiple DWD binaries can form within a 
single star particle. Specifically, we use an epanechnikov kernel where 
the kernel size is inversely proportional to the local density to assign 
the radial component of spherically symmetric offsets from the center of 
each star particle following \citet{Sanderson2020}.

\begin{figure}
	\includegraphics[width=\columnwidth]{figures/sfh_vs_fb.pdf}
    \caption{The metallicity-dependent binary fraction for close 
solar-type binaries with P $< 10^4$ days (black) plotted against the 
logarithm of $Z/Z_\odot$, with scatter points denoting the location of the 
metallicity grid used in our \cosmic\ simulations. The secondary axis 
shows the amount of mass in stars formed within each metallicity bin from 
galaxy \textbf{m12i} of the ``Latte" suite in the FIRE-2 simulations as a 
red histogram. The amount of stellar mass formed with super-solar 
metallicities dominates the distribution. Note that the primary axis shows 
$f_{\rm{b}}(Z)$ in linear scale, and the secondary axis shows amount of 
mass formed in log-scale. The opposing trends of these two distributions 
compete throughout this study.}
    \label{fig:sfh_vs_fb}
    \script{sfh_vs_fb.py}
\end{figure}

The metallicity-dependent close binary fraction for solar-type binaries 
($P_{\rm{orb}} < 10^4$ days), or simply the close binary fraction from 
here on, is shown in black in Figure \ref{fig:sfh_vs_fb} along with the 
mass in star particles from galaxy \textbf{m12i}, shown in red, as a 
function of our metallicity grid. The close binary fraction, $f_{\rm{b}}$, 
drops drastically across metallicity while the mass formed in 
\textbf{m12i} increases significantly. These two opposing trends compete 
throughout this study along with the impact of metallicity on single star 
evolution to form the final numerical distribution of systems in our DWD 
populations.
